Blame and Coercion: Together Again for the Second Time

### 2019-10-18 paper.tex
finish intro of sec4:framework

This section presents our second contribution, a framework for proving 
correctness of cast representations. The framework is a theorem saying that for 
all cast representation $ C $, if $ C $ satisfies a collection of properties, 
that an interpreter using $ C $ as the cast representation is extensionally 
equivalent to the standard interpreter for \lazyD\ cast calculus. 
We are working on a similar theorem for \lazyUD{}.

Firstly, let us define what we mean by \emph{cast representations}. The first 
three operators allow the interpreter to construct casts. We call them cast 
constructors. 
And the forth and last operator enable the interpreter to use casts.

\begin{definition}[Cast Representation]
	A cast representation is a set indexed by two types and has four interface 
	functions:
	\begin{description}
		\item[$ id(T) $] constructs an identity cast
		\item[$ seq(c_1,c_2) $] composes two casts
		\item[$ cast(T_1,l,T_2) $] constructs a cast from $ T_1 $ to $ T_2 $
		\item[$ applyCast(c,v) $] applies a cast onto a value
	\end{description}
\end{definition}

To use the framework, users needs to prove that their cast representation is 
\emph{surely LazyD}. The properties of surely \lazyD\ allow the framework to 
compute the result of $ applyCast(c,v) $ without referring to its actual 
definitions. This is because every casts at runtime is constructed by one of 
the three cast constructors. The first property handles casts constructed by $ 
id(T) $. And the second property handles $ seq(c,c) $. All remaining properties 
handles $ cast(T,l,T) $. They are effectively a duplication of the $ 
applyCast(c,v) $ in \ineffCEKD\ (Fig.~\ref{fig:applyCast-D-C}).

\begin{definition}[Surely \lazyD\ Cast Representation]
	A cast representation is surely \lazyD\ if
	\begin{enumerate}
		\item If $ v : T $, then $ applyCast(id(T), v) = \mathtt{succ} \; v $
		\item If $ \judgeType{v}{T_1} $,
		$ \judgeTypeFT{c_1}{T_1}{T_2} $, and
		$ \judgeTypeFT{c_2}{T_2}{T_3} $,\\
		then $ applyCast(seq(c_1,c_2),v) = 
		applyCast(c_1,v) >>= \lambda v.applyCast(c_2,v) $ \\
		where 
		\[
		\begin{array}{rcl}
		\rOOsucc{v} >>= f & = & f(v) \\
		\rOOfail{l} >>= f & = & \rOOfail{l}
		\end{array}
		\]
		\item If $ v : T_1 $ and $ \neg T_1 \smile T_2 $,
		then $ applyCast(cast(T_1, l, T_2),v) = \rOOfail{l} $
		\item If $ v : \star $, 
		then $ applyCast(cast(\TOOdyn,l,\TOOdyn),v) = \rOOsucc{v} $
		\item If $ v : P_1 $,
		then $ applyCast(cast(\star,l,P_2),\hcvOOinj{P_1}{v}) 
		= applyCast(cast(P_1,l,P_2),v) $
		\item If $ v : P $,
		then $ applyCast(cast(P,l,\star),v) = \rOOsucc{(\hcvOOinj{P}{v})} $
		\item If $ v : \POOunit $,
		then $ applyCast(cast(\POOunit,l,\POOunit),v) = \rOOsucc{v} $
		\item
		$ 
		applyCast(cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4}) ,
		\hcvOOfun{c_1}{\rho}{x}{b}{c_2}) \\
		= 
		\rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\rho}{x}{b}{seq(c_2,cast(T_2,l,T_4))})}$

@@

In subsection~\ref{ssec:framework:cek}, we introduce \effCEK{C}, a family 
of space-efficient semantics (in the style of CEK machines) indexed by cast 
representations ($ C $). 
In subsection~\ref{secc:framework:correctness}, we prove when $ C $ is 
$ surely \; \text{\lazyD} $, $ \judgeDeval{e}{o} $ if and only $ 
\judgeSeval{C}{e}{o} $, that is, every $ surely\;\text{\lazyD} $ cast 
representation respects the semantics of \lazyD\ cast calculus. Having this 
theorem means when people want to prove a \lazyD\ cast reprentation is correct, 
all they have to do is to prove that their cast representation is $ surely \;
LazyD $ and apply our framework to this proof, which should be much simpler 
than proving equivalence of semantics.

Our space-efficient CEK machine is parameterized over \emph{cast 
	representations}.

Some cast representations are monoids as well. They are useful in our 
correctness proof.

\begin{definition}[Monoid]
	A cast representation is a monoid if 
	for all $c : T_1 \Longrightarrow T_2$,
	$c_1 : T_1 \Longrightarrow T_2$,
	$c_2 : T_2 \Longrightarrow T_3$, and
	$c_3 : T_3 \Longrightarrow T_4$,
	\begin{enumerate}
		\item $seq(id(T_1),c) = c$,
		\item $seq(c,id(T_2)) = c$, and
		\item $seq(seq(c_1, c_2), c_3) = seq(c_1, seq(c_2, c_3))$.
	\end{enumerate}
\end{definition}

@@

	\caption{Space-efficient CEK machine $ \mathcal{S}(C) $}
	\label{machine-cekcc}
\end{figure}

\figref{machine-cekcc} defines the space-efficient CEK machine \effCEK{C}. This 
machine is similar to \ineffCEK{}. 
The most important differences are in values and continuations.
In \ineffCEK\, we have one value constructor for all casted value.
In \effCEK{C}, however, we do not have this generic value constructor, 
instead we push those casts into the value constructors most of the time and 
use $ \hcvOOinj{P}{v} $ to construct values casted to $ \TOOdyn $.
Non-casted values in \ineffCEK\ correspond to values in \effCEK{C} where the 
casts are identity (i.e. constructed by $ id(T) $).
In \ineffCEK\, a continuation have zero or more casts at the top.
In \effCEK{C}, however, every continuation has exactly one cast at the top.
Continuations in \ineffCEK\ that have no casts at the top correspond to 
continuations in \effCEK{C}\ whose casts are identity.
And continuations in \ineffCEK\ that have many casts at the top correspond to 
continuations in \effCEK{C}\ where those casts are composed by $ 
seq(c,c) $.

@@

\subsection{Surely \lazyD\ Cast Representations Respect \ineffCEKD}
\label{secc:framework:correctness}

@@

>>>>

\todo[inline]{What is the framework? Why can it simply proofs? Soundness?}

This section presents our second contribution, a framework for proving 
correctness of cast representations. The framework is a theorem saying that for 
all cast representation $ C $, if $ C $ satisfies a collection of properties, 
then instantiating our parameterized space-efficient interpreter with $ C $ 
gives an interpter that is extensionally equal to the standard interpreter for 
\lazyD\ cast calculus. We are working on a similar theorem for \lazyUD{}.

Firstly, let us define what we mean by \emph{cast representations} 
(Definition~\ref{def:cast-rep}). The first three operators allow the 
interpreter to construct casts. We call them \textit{cast constructors}. 
And the forth and last operator enable the interpreter to use casts.

\begin{definition}[Cast Representation]
	\label{def:cast-rep}
	A cast representation is a set indexed by two types and has four interface 
	functions:
	\begin{description}
		\item[$ id(T) $] constructs an identity cast
		\item[$ seq(c_1,c_2) $] composes two casts
		\item[$ cast(T_1,l,T_2) $] constructs a cast from $ T_1 $ to $ T_2 $
		\item[$ applyCast(c,v) $] applies a cast onto a value
	\end{description}
\end{definition}

To use the framework, users needs to prove that their cast representation is 
\emph{surely LazyD} (Definition~\ref{def:surely-lazyd}).
The definition confines the behavior of $ applyCast(c,v) $ so narrowly that it 
allows the framework to apply casts without referring to the actual definition 
of $ applyCast(c,v) $. This 
is because every casts at runtime is constructed by one of 
the three cast constructors. 
Property (1) handles casts constructed by $ id(T) $. 
And property (2) handles $ seq(c,c) $, which says applying the 
composition of $ c_1 $ and $ c_2 $ is the same as applying $ c_1 $ then $ c_2 $.
All remaining properties handles $ cast(T,l,T) $. They are effectively a 
duplication of the $ applyCast(c,v) $ in \ineffCEKD\ 
(Fig.~\ref{fig:applyCast-D-C}).

\begin{definition}[Surely \lazyD\ Cast Representation]
	\label{def:surely-lazyd}
	A cast representation is surely \lazyD\ if
	\begin{enumerate}
		\item If $ v : T $, then $ applyCast(id(T), v) = \mathtt{succ} \; v $
		\item If $ \judgeType{v}{T_1} $,
		$ \judgeTypeFT{c_1}{T_1}{T_2} $, and
		$ \judgeTypeFT{c_2}{T_2}{T_3} $,\\
		then $ applyCast(seq(c_1,c_2),v) = 
		applyCast(c_1,v) >>= \lambda v.applyCast(c_2,v) $ \\
		where 
		\[
		\begin{array}{rcl}
		\rOOsucc{v} >>= f & = & f(v) \\
		\rOOfail{l} >>= f & = & \rOOfail{l}
		\end{array}
		\]
		\item If $ v : T_1 $ and $ \neg T_1 \smile T_2 $,
		then $ applyCast(cast(T_1, l, T_2),v) = \rOOfail{l} $
		\item If $ v : \star $, 
		then $ applyCast(cast(\TOOdyn,l,\TOOdyn),v) = \rOOsucc{v} $
		\item If $ v : P_1 $,
		then $ applyCast(cast(\star,l,P_2),\hcvOOinj{P_1}{v}) 
		= applyCast(cast(P_1,l,P_2),v) $
		\item If $ v : P $,
		then $ applyCast(cast(P,l,\star),v) = \rOOsucc{(\hcvOOinj{P}{v})} $
		\item If $ v : \POOunit $,
		then $ applyCast(cast(\POOunit,l,\POOunit),v) = \rOOsucc{v} $
		\item
		$ 
		applyCast(cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4}) ,
		\hcvOOfun{c_1}{\rho}{x}{b}{c_2}) \\
		= 
		\rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\rho}{x}{b}{seq(c_2,cast(T_2,l,T_4))})}$

@@

If a cast representation is surely LazyD, proving this should be easier than 
proving equivalence to the standard semantics of cast calculus.

Some cast representations are also \textit{monoids}. They are useful in the 
proof of our framework, and might be interesting elsewhere.

\begin{definition}[Monoid]
	A cast representation is a monoid if 
	for all
	$c_1 : T_1 \Longrightarrow T_2$,
	$c_2 : T_2 \Longrightarrow T_3$, and
	$c_3 : T_3 \Longrightarrow T_4$,
	\begin{enumerate}
		\item $seq(id(T_1),c_1) = c_1$,
		\item $seq(c_1,id(T_2)) = c_1$, and
		\item $seq(seq(c_1, c_2), c_3) = seq(c_1, seq(c_2, c_3))$.
	\end{enumerate}
\end{definition}

In subsection~\ref{ssec:framework:cek}, we introduce \effCEK{C}, a family 
of space-efficient semantics indexed by cast representations ($ C $). 
In subsection~\ref{secc:framework:monoid-correct}, we prove when $ C $ is 
$ surely \; \text{\lazyD} $ and is a monoid as well, $ \judgeDeval{e}{o} $ if 
and only if $ \judgeSeval{C}{e}{o} $. In 
subsection~\ref{secc:framework:all-correct}, we strengthen the theorem to drop 
the requirement that $ C $ must be a monoid, that is, every surely \lazyD\ cast 
representation is correct.

@@

	\caption{Space-efficient CEK machine $ \mathcal{S}(C) $}
	\label{machine-cekcc}
\end{figure}
 (in the style of CEK machines)
\figref{machine-cekcc} defines the space-efficient CEK machine \effCEK{C}. This 
machine is similar to \ineffCEK{}. 
The most important differences are in values and continuations.
In \ineffCEK\, we have one value constructor for all casted value.
In \effCEK{C}, however, we do not have this generic value constructor, 
instead we push those casts into the value constructors most of the time and 
use $ \hcvOOinj{P}{v} $ to construct values casted to $ \TOOdyn $.
Non-casted values in \ineffCEK\ correspond to values in \effCEK{C} where the 
casts are identity (i.e. constructed by $ id(T) $).
In \ineffCEK\, a continuation have zero or more casts at the top.
In \effCEK{C}, however, every continuation has exactly one cast at the top.
Continuations in \ineffCEK\ that have no casts at the top correspond to 
continuations in \effCEK{C}\ whose casts are identity.
And continuations in \ineffCEK\ that have many casts at the top correspond to 
continuations in \effCEK{C}\ where those casts are composed by $ 
seq(c,c) $.

@@

\subsection{Surely \lazyD\ Cast Representations Respect \ineffCEKD\ if They are 
Monoids}
\label{secc:framework:monoid-correct}

@@

\subsection{Surely \lazyD\ Cast Representations Respect \ineffCEKD}
\label{secc:framework:all-correct}
